Framing active school travel in Ontario, or how spinach is good for you

### 2023-05-26 editorial-materials/YTRA-D-22-01243-response-to-reviewers.Rmd

---
title: "Response to reviewers"
output: pdf_document
link-citations: yes
linkcolor: blue
bibliography: 
  - '`r system("kpsewhich ../AST-Framing-Ontario/bibliography.bib", intern=TRUE)`'
---

@@

We wish to thank both reviewers for their constructive feedback. We feel that the paper has been much improved as a result of responding to their comments.

In particular, we recognize that the methodology we chose is perhaps infrequently applied in transport research. That said, the purpose of the paper was to examine the way policy with respect to AST are communicated in Canada's most populous province. As the methodology is not common in transportation, we tried to explain more of it, which may be why the paper came off as methologically-focused. In this revision, we aimed to strike a balance between the focus (how is AST discussed in policy) and ensuring that readers understood the possibly unfamiliar methods.

@@

Finally, we would like to close with a brief note on the methodology. Natural language processing and analysis of frames are still not widely adopted in transportation research. Based on our experience, we tend to see computer-assisted text analysis as a complement rather than a substitute to expert reading of the documents. We would not recommend to proceed to interpretation without reading the documents, but the supervised automation helps to detect patterns that can be validated by an expert based on spot checks or more in-depth reading of the original sources. Like with any qualitative research, interpretation of the frames may differ by expert; in this research we aimed to develop a consensus interpretation by involving all authors in the discussion of the results. Furthermore, the research was conducted as an open, reproducible project: all code and data are available in a public repository^[\url{https://github.com/desjae/AST-Framing-Analysis-Ontario}], which means that all our assumptions are open to scrutiny. Moreover, other researchers can refer to the repository to expand or challenge our conclusions.

### Comment 4 {#r1-comment-4}

> This brings me to my primary concern - I couldn't help but wonder if this piece might be a better fit to TR-B or C - given the emphasis on demonstrating method/tools - and the production of conclusions that are rather unsurprising - the last sentence of the paper alone is something that has been the focus of much work, in practice, for decades.

Thank you for sharing this concern. The methods are not really new outside of their application to this research. Natural language processing is a field mature enough to have its own textbooks; framing analysis is also widespread, although its use in transportation research remains limited. We do not think of either method as emerging technologies or sufficiently innovative to make this a "methodological" paper. Instead, what we think is new is the use of these methods to reveal something interesting about the effectiveness of transportation policy.

@@

### Comment 2 {#r2-comment-2}

> However in my opinion is not within the scope of Transportation Research Part A. I did not understand the link with the policy and practice.

We wish to reiterate our response to [Reviewer 1](#r1-comment-4). We part from the working hypothesis that if the Editor, having perused the paper, decided to send it out for peer-review, they must have decided that the paper is within the scope of the journal.

Having said that, we made a renewed effort to clarify the link of our research to policy and practice. In particular, we think Reviewer 1 is right on target with their [comment](#r1-comment-5): despite the mass of academic research in transportation, we do not seem to be further along in effectively communicating with the public the ever higher stakes of our ongoing love affair with the car. The problem does not seem to be the lack of policies, but awareness and buy-in by the public. Why is this the case when it comes to active school travel?

@@

The paper has been reworked to (hopefully) improve the link between framing and policy. The argument goes as follows:

1. Public policies depend on what we collective believe is possible [@beland2014developing, p. 551; also see @beland2008nationalism; @akerlof2010identity]

2. Frames are essential for defining the parameters of our collective beliefs: frames can activate or restrict particular responses in the intended audience [@panFramingAnalysisApproach1993] by positioning existing solutions as suitable to issues. This can limit awareness of other policy approaches [@mahFramecriticalPolicyAnalysis2014]. Contrariwise, they can expand the potential of policy by revealing more diverse options than conventional wisdom would have [@bosomworth2015climate].

@@

4. Active school travel is framed in Ontario in such a way that it downloads the responsibility for action to individuals; it does not make a convincing call for collective action. It is also fairly neutral at a time when some stakes (e.g., climate change) are extremely high. By adopting an empirical-scientific approach, it does not appeal to moral reasons for change. And it ends up being conservative and anodyne.

### Comment 3 {#r2-comment-3}

> The research finds that a specific policy is not working as intended and it seems to be very expensive. How can this result be generalised and used to say if any policy is "good value for money"? The authors need to convince me about the practical application of their method. Can you identify clear policy recommendations that can be used in practice?

Point 4 [above](#r2-comment-2) highlights some things that we learn from this research, and suggests some ways in which frames around AST could be changed.

### 2023-05-26 AST-Framing-Ontario/AST-Framing-Ontario.Rmd
Resubmission to TR-A after revisions by ED, DC, DR, EOW, and AP

---
title: "Framing active school travel in Ontario: , or how spinach is good for you"
author:
  - name: Elise Desjardins
    email: desjae@mcmaster.ca
    affiliation: McMaster University
  - name: Jason Lam
    email: lamj54@mcmaster.ca
    affiliation: McMaster University
  - name: Darcy Reynard 
    email: reynard@ualberta.ca
    affiliation: University of Alberta
  - name: Damian Collins
    email: damian.collins@ualberta.ca
    affiliation: University of Alberta
  - name: E. Owen D. Waygood
    email: owen.waygood@polymtl.ca
    affiliation: Polytechnique Montreal
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    correspondingauthor: true
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, 1280 Main St. W, Hamilton, Ontario, Canada L8S 4K1
  - code: University of Alberta
    address: Department of Earth and Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta, Canada T6G 2E3
  - code: Polytechnique Montreal
    address: Department of Civil, Geological and Mining Engineering, 2500, chemin de Polytechnique Montreal, Montreal, Quebec, Canada H3T 1J4
footnote:
  - code: 1
    text: "Corresponding author."
abstract: |
  Active school travel (AST) is a relevant policy issue in Ontario, Canada, as reflected in provincial support for school travel planning (STP) efforts. Two pillars of school travel plans are education and encouragement. For these pillars to stand, all relevant stakeholders must be cognizant of and understand well the issues and stakes. For this reason, how AST is communicated to them is essential. In this research we adopt frame analysis to investigate the way various organizations communicate around AST to the public. A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and helps to set the parameters for conversations about policy. In the case of AST, framing can influence what policy alternatives are perceived as available by children, parents, and their wider communities. To investigate framing of AST in Ontario we use natural language processing techniques to analyze publicly available documents from Ontario stakeholders involved in school travel planning. We then compared the findings from these documents to a selection of academic studies on AST. We conclude that framing of AST in Ontario is mostly empirical-scientific in style, and largely in agreement with academic research on AST. On the other hand, it it does not appeal to moral reasons for change, and is therefore oddly soothing and lacking in urgency. Furthermore, the frames used download responsibility for change to households, which limits the scope of policy alternatives by keeping collective action out of the frame.

@@

Walking and bicycling to school, commonly known as active school travel (AST), have declined in North America for decades [@rothmanDeclineActiveSchool2018], but the appetite for these modes is beginning to rise in many Canadian communities. For instance, research in Toronto, Ontario found that 40% of children who were driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. Another study in London, Ontario, reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public policy perspective, increasing active travel to school represents a major opportunity to improve children's health and wellbeing [@faulknerActiveSchoolTransport2009], to mitigate traffic and improve perceptions of safety [@rothmanAssociationsParentsPerception2015], as well as to reduce carbon and other noxious emissions [@bearman2014modelling; @goodman2019scenarios].

To promote this trend, in 2017 the Government of Ontario began to support communities develop AST initiatives through a program called _Ontario Active School Travel_ (OAST). By 2021, the program, run by _Green Communities Canada_ had awarded over CAD 2 million and provided resources to over 25 projects across Ontario [@GreenCommunities2016]. A popular intervention in many communities is school travel planning (STP) [see @GreenCommunitiesProjects]. In this context, STPs are "school-specific" interventions led by a facilitator who works with a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST and implementing solutions or activities to make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

How AST is presented to a target audience (e.g., parents or the general public) is essential to define the parameters of what is possible policy-wise. Drawing from concepts in communication studies, the objective of this paper is to investigate how AST is framed in Ontario. Gamson and Modigliani [-@gamsonChangingCulture1987] define a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. From this perspective, stakeholders involved in STP efforts play an important role in shaping public perception about AST by generating frames that  create meaning, thus influencing the degree of attention and recognition of a "problem" that needs to be addressed through behaviour change or new policies.

Our research is supported by natural language processing techniques, applied to the analysis of documents produced by three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia). We assemble a corpus of texts from public sources that present information for the general public or parents who are interested in AST and examine word frequency, bigrams, and concordances in these selected documents. Furthermore, we use topic analysis to identify themes presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the scholarly literature on AST and materials shared with the public. The results of this analysis, along with spot checks of the corpus by the authors, lead us to conclude that AST if framed as a matter of individual choices instead of collective action. The message is reasonable but anodyne, and it fails to communicate a sense of urgency about health and environmental issues, which may help to explain the limited success of policy.

@@

<!--Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of household, social, neighbourhood, and policy environments. The SEM helps to identify multiple determinants that need to be addressed by school-based interventions to increase AST, with the consensus being that interventions should target multiple levels [@mitraIndependentMobilityMode2013]. -->

Rather than describing an extensive list of factors that influence AST and mode choice for travel to school in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we focus on two key aspects that may be targeted from a policy perspective: perceptions and behaviors by parents and built environmental factors.

@@

First, research has found that parents are key decision-makers with respect to children's form of travel to school [@rothmanAssociationsParentsPerception2015]. Particularly, parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010; @mandic2020differences] and their children's skills [@mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010] tend to influence mode choice to school. Furthermore, children's mode choice to school is strongly influenced by their parents' travel behaviours, the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as daily parental support [@mahDoesParentalSupport2017a]. This suggests that shifting parental perceptions and habits is important.

With respect to the built environment, distance between home and school is most negatively associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012]: ironically, concerns about traffic become a self-fulfilling prophecy.

@@

AST has become a policy issue on the education and public health agendas in Ontario, supported, albeit in a modest way, by financial contributions from the provincial government. Still, the support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda. One of the main vehicles to promote AST is school travel planning (STP), which has been implemented in Canada since at least the late 2000s. Within the STP process, facilitators establish multi-sector committees which intervene at the participating school through a range of activities consisting of _[E]ducation_ strategies, _[E]ncouragement_ through in-person events or programs, _[E]ngineering_ improvements to or around the school site, and _[E]nforcement_ of traffic speeds around schools [or 4Es, @langUnderstandingModalChoice2011; @mammenActiveSchoolTravel2014]. Following the first large-scale evaluation of STP as an intervention in Canada, AST rates increased and 14% of parents reported a perceived reduction in traffic outside the school [@buliungSchoolTravelPlanning2011]. Assessments of the efficacy of STP [@buttazzoniPromotingActiveSchool2019; @mammenActiveSchoolTravel2014] indicate its potential to encourage behaviour change and adoption of AST.

The relevance of policy frames is highlighted by @beland2014developing, who remarks "framing processes as public relations are central to policy issues related to sustainable transportation". Seeing how highly influential parental perceptions are to AST, it is important to understand AST-related communications, since the construction of frames for policy positions or public issues can activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Besides parents and children, framing can also affect broader policy support for AST in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortia representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In this sense, framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo. Contrariwise, framing can be instrumental to revealing more diverse policy options than more diverse policy options than conventional wisdom would have [@bosomworth2015climate]. Ultimately, it is important to understand framing of policy issues because creating meaning through frames contributes to the constant process of shaping of our collective identities, uncovering what we collectively find (un)acceptable and why [@beland2014developing].

Given the heavy focus of STP activities on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], it is important for the issues to be effectively framed. In the past, parents have been found to express different understandings, language, and perceptions than planners of how the built environment can influence school travel: while parents tend to view mixed land use as conducive for driving, transport planners see mixed uses as a key for encouraging more active travel [@buliungLivingJourneySchool2021]. Similarly mismatched understandings have been detected with respect to factors like the convenience of different modes to school [@langUnderstandingModalChoice2011]. As a consequence, STP stakeholders must pay special attention to parents' and communities' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions [@buttazzoniSupportingActiveSchool2018].

So framing is important, but not all frames are equally effective. Frames can appeal to economic reasons, to science and logic, or to moral obligation. For example, @severson2015moral report that overall support for climate change policy is increased more when scientific frames, secular moral frames, and economic equity frames are used, compared to economic efficiency or religious moral frames. Frames can also be positive or negative, and there is evidence that negative framing may be more effective at motivating individuals, for example, to change transport modes to reduce CO2 emissions [@waygoodCO2ValenceFraming2018]. As well, the effectiveness of the framing depends on the psychological distance to the relevant issues: how socially, temporally, geographically, or hypothetically close an individual perceives themselves to the issues can affect their support for various policies [@maiella2020psychological]. For this reason, stakeholders involved in STP must make conscious choices about how issues surrounding AST are communicated if the goal is to increase support for and buy-in of AST policies.

@@

We analyzed word and document frequency for each corpus of text. Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. Policy documents and academic papers reference _active_, _travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpus also has _safety_ and _traffic_ as common words which suggests that these are common concerns for parents. The word _physical_ is present in each corpus, but this could refers to _physical activity_ or the _physical environment_. Furthermore, documents from STP stakeholder groups discuss _resources_, _information_, and _services_ about school travel. Unlike the academic papers, policy+practice documents frequently include the words _route_ or _routes_. This could indicate the role of STP stakeholder groups in identifying safe routes to school to share with parents or families. In the section below, the context in which these terms are used is explored further.

@@

Bigrams refer to a pair of consecutive words. We combined all municipality, school board, and transportation consortia documents into the policy+practice corpus. Figure \ref{fig:policy-visual} shows all of the bigrams that occur more than 10 times in this corpus. This figure highlights the main ideas that are presented to the public in each of the policy corpora. The directional arrows indicate the arrangement of the words (e.g., active travel and not travel active) and the colour gradient of the arrows corresponds to the most frequently mentioned pairs (e.g., bigrams with darker arrows are found in the corpus more often).

STP stakeholders discuss _physical activity_ and _public health_ in the context of AST. In addition, _travel planning_, _bike lanes_, and _safe routes_ are also identified, conceivably as either proposed solutions or built environment factors that support AST. Key issues related to transport such as _traffic safety_, _air quality_, and _greenhouse gases_ are conveyed to the public through these policy documents. It is not surprising to find this focus given that municipalities in Ontario are concerned about climate change and have increasingly looked to active travel to offset transport-related emissions in urban areas. We also found _mental health_, _walk zone_, and _green communities_ as common pairs of consecutive words. Their presence makes sense given the involvment of  _Green Communities Canada_ in AST initiatives in Ontario. Overall, the policy documents from STP stakeholder groups seem to focus on four key areas: i) benefits or impacts of AST; ii) mechanisms of intervention; iii) concerns or barriers; and iv) supports for AST. This interpretation indicates that the general public accessing information about AST in Ontario is informed about a wide range of content related to this issue.

Next, we analyzed bigrams in the academic corpus separately to comparE with the policy+practice corpus. Figure \ref{fig:academic-visual} indicates that academic papers include several common bigrams that were also found in the policy documents including _physical activity_, which is the top bigram, as well as _traffic safety_ and _safe routes_. However, many other factors are identified in the academic corpus that are not presented to the general public through policy documents. Terms like _built environment_, _independent mobility_, and _urban form_ are other frequent pairs of consecutive words. Academic papers also often discuss _distance home_, _car ownership_, _household income_, and _population density_, which are factors that have been found to influence AST. Finally, the presence of _statistically significant_ among the top bigrams indicates that researchers often aim to identify associations using statistical measures. We found that the academic corpus focuses on a greater range of topics than found in the policy+practice documents.

@@

The academic corpus has a higher number of topics likely due to the volume of papers included. The following topics were identified based on the clusters of words: (1) physical activity; (2) safe routes to school for active travel; (3) behaviours and attitudes; (4) bicycling; (5) walking; (6) built environment; (7) trip choice; (8) active school travel interventions; (9) parents; (10) children's independent mobility; and (11) public health and policies. This corpus reflects a broader range of topics than the policy+practice corpus.

@@

The secondary frame in policy documents is that AST is accessible and feasible for children and parents. Despite the emphasis on the logistics of busing in the topic modelling, the documents indicate that there is an opportunity for or prospect of behaviour change. Some cities and schools explain how children and parents can leave the car at home and make the journey to school on foot or by bike. This frame encourages the public to evaluate their own travel decisions and to access resources (e.g., walking skills checklist) that will help them make AST a first choice. Examples of this secondary frame include:

@@

Finally, both AST frames present the benefits of AST. STP stakeholder groups seem to be communicating that AST is a healthy and accessible option as a result of school travel plans and improvements to the built environment. This could include examples of various efforts that are underway to support AST including route planning.

>>>>

---
title: "Framing active school travel in Ontario, or how spinach is good for you"
author:
  - name: Elise Desjardins
    email: desjae@mcmaster.ca
    affiliation: McMaster University
  - name: Jason Lam
    email: lamj54@mcmaster.ca
    affiliation: McMaster University
  - name: Darcy Reynard 
    email: reynard@ualberta.ca
    affiliation: University of Alberta
  - name: Damian Collins
    email: damian.collins@ualberta.ca
    affiliation: University of Alberta
  - name: E. Owen D. Waygood
    email: owen.waygood@polymtl.ca
    affiliation: Polytechnique Montreal
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    correspondingauthor: true
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, 1280 Main St. W, Hamilton, Ontario, Canada L8S 4K1
  - code: University of Alberta
    address: Department of Earth and Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta, Canada T6G 2E3
  - code: Polytechnique Montreal
    address: Department of Civil, Geological and Mining Engineering, 2500, chemin de Polytechnique Montreal, Montreal, Quebec, Canada H3T 1J4
footnote:
  - code: 1
    text: "Corresponding author."
abstract: |
  Active school travel (AST) is promoted in many jurisdictions, including Ontario, Canada, where there is provincial support for school travel planning (STP) efforts. Two pillars of school travel plans in the province are education and encouragement. For these pillars to stand, relevant stakeholders must be cognizant of and understand the issues and stakes. For this reason, we deem it important to understand how AST is communicated. In this research we adopt framing analysis to investigate the ways in which various organizations communicate to the public around AST. A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and thus helps to set the parameters for conversations about policy. In the case of AST, framing can influence what policy alternatives are perceived as available by children, parents, and their wider communities. The research is supported by natural language processing techniques applied to publicly available documents from Ontario stakeholders involved in school travel planning. We then compare the findings from these documents to a selection of academic studies on AST. We conclude that framing of AST in Ontario is mostly empirical-scientific in style, and largely in agreement with academic research on AST. However, it is essentially conservative and does challenge to the status quo of motorized travel. Furthermore, the frames tend to download responsibility for change to households, which limits the scope of policy alternatives by keeping collective action out of the frame.

@@

Walking and bicycling to school, commonly known as active school travel (AST), have declined in North America for decades [@rothmanDeclineActiveSchool2018], but the appetite for these modes is beginning to rise in many Canadian communities. For instance, research in Toronto, Ontario found that 40% of children who were driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. Another study in London, Ontario, reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public policy perspective, increasing active travel to school presents major opportunities to improve children's health and wellbeing [@faulknerActiveSchoolTransport2009], to mitigate traffic and improve perceptions of safety [@rothmanAssociationsParentsPerception2015], as well as to reduce emissions and other forms of pollution [@bearman2014modelling; @goodman2019scenarios].

To promote this trend, in 2017 the Government of Ontario began to support communities develop AST initiatives through a program called _Ontario Active School Travel_ (OAST). By 2021, the program, run by _Green Communities Canada_ had awarded over CAD 2 million and provided resources to over 25 projects across Ontario [@GreenCommunities2016]. A popular intervention in many communities is school travel planning (STP) [see @GreenCommunitiesProjects]. In this context, STPs are "school-specific" interventions led by a facilitator who works with a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST and implementing solutions or activities to make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@hinckson2018school].

How AST is presented to a target audience (e.g., parents or the general public) is essential to define the parameters of what is possible policy-wise. Drawing from concepts in communication studies, the objective of this paper is to investigate how AST is framed in Ontario. Gamson and Modigliani [-@gamsonChangingCulture1987] define a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. From this perspective, stakeholders involved in STP efforts play an important role in shaping public perceptions about AST by generating frames that create meaning, and thus influence the degree of attention and recognition of a "problem" that needs to be addressed through behavior change or new policies.

Our research is supported by natural language processing techniques, applied to the analysis of documents produced by three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia). We assemble a corpus of texts from public sources that present information for the general public or parents who are interested in AST and examine word frequency, bigrams, and concordances in these selected documents. Furthermore, we use topic analysis to identify themes presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the scholarly literature on AST and materials shared with the public. The results of this analysis, along with manual checks of the corpus by the authors, lead us to conclude that AST is framed as a matter of individual choices instead of collective action. The message is reasonable but inherently conservative and it fails to communicate a sense of urgency about health and environmental issues, which may help to explain the modest success of related policies [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014; @mammenActiveSchoolTravel2014]

@@

<!--
ED:
Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behavior is understood within the context of household, social, neighbourhood, and policy environments. The SEM helps to identify multiple determinants that need to be addressed by school-based interventions to increase AST, with the consensus being that interventions should target multiple levels [@mitraIndependentMobilityMode2013].

AP:
I removed the SEM from the paper because it was used mainly to appeal to the social science and medicine crowd, but does not seem to be essential for the presentation to the transportation research community
-->

Rather than describing an extensive list of factors that influence AST and mode choice for travel to school in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we focus on two key aspects that may be targeted from a policy perspective: built environmental factors and perceptions and behaviors of parents.

The built environment is a key factor for physical planning, and in this respect distance between home and school is the factor most negatively associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012]: ironically, concerns about the dangers of traffic become a vicious cycle [@rothmanSchoolEnvironmentStudent2017].

@@

Secondly, research has found that parents are key decision-makers with respect to children's mode of travel to school [@rothmanAssociationsParentsPerception2015]. Particularly, after distance [@curtis2015built], parental perceptions of the built or school environment, as well as of their children's skills, have been found to be key in influencing mode choice to school [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010; @mandic2020differences; @mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010]. For example, @ramanathanHappinessMotionEmotions2014a found that parents and children who perceived active travel as beneficial for health and wellbeing benefits were more likely to use active modes to get to school. Furthermore, children's mode choice to school is strongly influenced by their parents' travel behaviors, the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as daily parental support [@mahDoesParentalSupport2017a]. This suggests that shifting parental perceptions and habits is important.

@@

AST has become a policy issue on the education and public health agendas in Ontario, and is supported, albeit in a modest way, by financial contributions from the provincial government. Still, endorsement from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda. One of the main vehicles to promote AST is school travel planning (STP), which has been implemented in Canada since at least the late 2000s. Within the STP process, facilitators establish multi-sector committees which intervene at the participating school through four pillars consisting of _[E]ducation_ strategies, _[E]ncouragement_ through in-person events or programs, _[E]ngineering_ improvements to or around the school site, and _[E]nforcement_ of traffic speeds around schools [or 4Es, @langUnderstandingModalChoice2011; @mammenActiveSchoolTravel2014]. Of the 4Es, [E]ngineering targets the first set of factors discussed in the preceding section (the built environment), whereas [E]ducation and [E]ncouragement target perceptions.

The relevance of policy frames with respect to education and encouragement is highlighted by @beland2014developing, who remarks "framing processes as public relations are central to policy issues related to sustainable transportation". Seeing how influential parental perceptions are to AST, it is important to understand AST-related communications, since the construction of frames for policy positions or public issues can activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Besides parents and children, framing can also affect broader policy support for AST in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortia representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In this sense, framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo. Contrariwise, framing can be instrumental to revealing more diverse policy options than conventional wisdom would have [@bosomworth2015climate]. Ultimately, it is important to understand framing of policy issues because creating meaning through frames contributes to the constant process of shaping of our collective understanding of issues, uncovering what we collectively find (un)acceptable and why [@beland2014developing].

Given the role of education and encouragement in STP, it is important for the issues to be effectively framed. In the past, parents have been found to express understandings, language, and perceptions regarding the influence of the built environment on school travel that markedly differ from those of planners: while parents tend to view mixed land use as conducive for driving, transport planners see mixed uses as a key for encouraging more active travel [@buliungLivingJourneySchool2021]. Similarly mismatched understandings have been detected with respect to factors like the convenience of different modes to school [@langUnderstandingModalChoice2011]. As a consequence, STP stakeholders must pay special attention to parents' and communities' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions [@buttazzoniSupportingActiveSchool2018].

Framing is important, but not all frames are equally effective. Frames can appeal to economic reasons, to science and logic, or to moral obligation. For example, @severson2015moral report that overall support for climate change policy increases more when scientific frames, secular moral frames, and economic equity frames are used, compared to economic efficiency or religious moral frames. Frames can also be positive or negative, and there is evidence that negative framing may be more effective at motivating individuals, for example, to change transport modes to reduce CO2 emissions [@waygoodCO2ValenceFraming2018]. As well, the effectiveness of the framing depends on the psychological distance to the relevant issues: how socially, temporally, geographically, or hypothetically close an individual perceives themselves to the issues can affect their support for various policies [@maiella2020psychological]. For this reason, stakeholders involved in STP must make conscious choices about how issues surrounding AST are communicated if the goal is to increase support for and buy-in of AST policies.

@@

We analyzed word and document frequency for each corpus of text. Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. Policy documents and academic papers reference _active_, _travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpus also has _safety_ and _traffic_ as common words, which suggests that these are common concerns. The word _physical_ is present in each corpus, but this could refers to _physical activity_ or the _physical environment_. Furthermore, documents from STP stakeholder groups discuss _resources_, _information_, and _services_ about school travel. Unlike the academic papers, policy+practice documents frequently include the words _route_ or _routes_. This could indicate the role of STP stakeholder groups in identifying safe routes to school to share with parents or families. In the section below, the context in which these terms are used is explored further.

@@

Bigrams refer to pairs of consecutive words. We combined all municipality, school board, and transportation consortia documents into the policy+practice corpus. Figure \ref{fig:policy-visual} shows all of the bigrams that occur more than 10 times in this corpus. This figure highlights the main ideas that are presented to the public in each of the policy corpora. The directional arrows indicate the arrangement of the words (e.g., active travel and not travel active) and the colour gradient of the arrows corresponds to the most frequently mentioned pairs (e.g., bigrams with darker arrows are found in the corpus more often).

STP stakeholders discuss _physical activity_ and _public health_ in the context of AST. In addition, _travel planning_, _bike lanes_, and _safe routes_ are also identified, conceivably as either proposed solutions or built environment factors that support AST. Key issues related to transport such as _traffic safety_, _air quality_, and _greenhouse gases_ are conveyed to the public through these policy documents. It is not surprising to find this focus given that municipalities in Ontario are concerned about climate change and have increasingly looked to active travel to offset transport-related emissions in urban areas. We also found _mental health_, _walk zone_, and _green communities_ as common pairs of consecutive words. The latter makes sense given the involvment of  _Green Communities Canada_ in AST initiatives in Ontario. Overall, the policy documents from STP stakeholder groups seem to focus on four key areas: i) benefits or impacts of AST; ii) mechanisms of intervention; iii) concerns or barriers; and iv) supports for AST. This interpretation indicates that the general public accessing information about AST in Ontario is informed about a wide range of content related to this issue.

Next, we analyzed bigrams in the academic corpus separately to compare with the policy+practice corpus. Figure \ref{fig:academic-visual} indicates that academic papers include several common bigrams that were also found in the policy documents including _physical activity_, which is the top bigram, as well as _traffic safety_ and _safe routes_. However, many other factors are identified in the academic corpus that are not presented to the general public through policy documents. Terms like _built environment_, _independent mobility_, and _urban form_ are examples of these factors. Academic papers also often discuss _distance home_, _car ownership_, _household income_, and _population density_, which are factors that have been found to influence AST. Finally, the presence of _statistically significant_ among the top bigrams indicates that researchers often aim to identify associations using statistical measures. We found that the academic corpus focuses on a greater range of topics than found in the policy+practice documents.

@@

The academic corpus has a higher number of topics likely due to the volume of papers included. The following topics were identified based on the clusters of words: (1) physical activity; (2) safe routes to school for active travel; (3) behaviors and attitudes; (4) bicycling; (5) walking; (6) built environment; (7) trip choice; (8) active school travel interventions; (9) parents; (10) children's independent mobility; and (11) public health and policies. This corpus reflects a broader range of topics than the policy+practice corpus.

@@

The secondary frame in policy documents is that AST is accessible and feasible for children and parents. Despite the emphasis on the logistics of busing in the topic modelling, the documents indicate that there is an opportunity for or prospect of behavior change. Some cities and schools explain how children and parents can leave the car at home and make the journey to school on foot or by bike. This frame encourages the public to evaluate their own travel decisions and to access resources (e.g., walking skills checklist) that will help them make AST a first choice. Examples of this secondary frame include:

@@

Finally, both frames present the benefits of AST. STP stakeholder groups seem to be communicating that AST is a healthy and accessible option as a result of school travel plans and improvements to the built environment. This could include examples of various efforts that are underway to support AST including route planning.
